\documentclass{article}
\usepackage{graphicx} % Required for inserting images
\usepackage[top=0.9in, bottom=1in, left=1.5in, right=1.5in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[icelandic]{babel}
\usepackage[T1]{fontenc}
\usepackage[sc]{mathpazo}
\usepackage[parfill]{parskip}
\renewcommand{\baselinestretch}{1.2}
\usepackage{booktabs,tabularx}
\usepackage{multirow}
\usepackage{enumerate}
\usepackage{adjustbox}
\usepackage{multicol}
\usepackage{xcolor}
\usepackage{algpseudocode}
\usepackage{tikz}
\usepackage{nicefrac}
\usepackage{changepage}
\usetikzlibrary{arrows, positioning, calc, graphs}
\usepackage{amsmath, amsfonts, amssymb, amsthm}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{minted}
\usemintedstyle{manni}
\title{Formal Languages and Computability 8}
\author{Ragnar Björn Ingvarsson, rbi3}
\tikzset{->, >=stealth', shorten >=1pt, node distance=2cm,thick, main node/.style={circle,draw,minimum size=3em}}

\begin{document}
\renewcommand\thepage{}

	\maketitle

	\newpage
	\setcounter{page}{1}
	\renewcommand\thepage{\arabic{page}}

	\section{}
	\begin{itemize}
		\item[a)] We see that putting $1011$ into the DFA leads us into 
			the reject state $q_r$ so the string is not in the language 
			that $M$ describes.
		\item[b)] Here, since the string $1011$ is not a string accepted by 
			$M$, $\langle M,1011\rangle$ is not in $A_{DFA}$, since $A_{DFA}$ 
			requires that $M$ accepts the string.
		\item[c)] Here we see that $1010$ is accepted by $M$ so 
			$\langle M,1010\rangle \in A_{DFA}$.
		\item[d)] This is not true since $EQ_{DFA}$ requires two DFA's, 
			$\langle M,N\rangle$ and so with only $\langle M\rangle$ we 
			cannot compare two DFA's since only one is given.
	\end{itemize}

	\section{}
	To prove that $D$ is countable, we can make do with the fact that since 
	$\mathbb{N}$ is countable, we only need to show that there exists an 
	injection $D\rightarrow \mathbb{N}$ since if that is the case, the 
	the cardinality of $D$, $|D| \leq |\mathbb{N}|$, meaning that since 
	$\mathbb{N}$ is countable, $D$ also has to be countable.

	We then let $f: \mathbb{N}\times\mathbb{N}\times\mathbb{N}\rightarrow\mathbb{N}$ be defined 
	as $f(n,m,o) = 2^n\cdot3^m\cdot5^o$, $n,m,o\in\mathbb{N}$. To prove that it is injective, we 
	assume that for some $a,b,c,m,n,o\in\mathbb{N}$,
	\[
		f(a,b,c) = f(n,m,o)
	\]
	\[
		\iff 2^a\cdot3^b\cdot5^c = 2^n\cdot3^m\cdot5^o
	\]
    which clearly shows that $a=n$, $b=m$, $c=o$. Therefore, $f: D\rightarrow\mathbb{N}$ is injective so $D$ is countable.

	\section{}
	\begin{itemize}
		\item[a)] Since we are working with a set of binary strings, the 
			set $A$ is countably infinite as $A$ is a subset of the set 
			of all possible binary strings, according to the Note on 
			Lemma 53 in lecture 8. If we would be working with binary 
			sequences however, the set $A$ would be uncountable.
		\item[b)] Since $w = 1.010101010\dots$ is an infinite binary 
			sequence, it is not in $A$ since $A$ only accepts finite binary 
			strings.
		\item[c)] $A$ is decidable since for any string $w$ which is 
			finite according to our definition in a), we can calculate 
			$\sqrt{8}$ to the precision of $w$ and then compare them, 
			all in finite time, then reject if $w>\sqrt{8}$ and accept 
			otherwise.
	\end{itemize}

	\section{}
    Here we will construct a new DFA, $C$, that accepts only strings that are 
	accepted by S and not by R. We then see that if $L(S)\subseteq L(R)$ 
	then $L(C) = \varnothing$. 

	Lets assume a TM $F = F(\langle R,S\rangle)$ which constructs a DFA $C$ 
	as we described. Then we can run a TM $T$ that decides on $E_{DFA}$ on 
	$C$. If $T$ accepts, $F$ accepts, if not, it rejects.

	TM $F$ then accepts if $L(C) = \varnothing$ and rejects otherwise so it 
	decides on the language $A$.
\end{document}
