\documentclass{article}
\usepackage{graphicx} % Required for inserting images
\usepackage[top=0.9in, bottom=1in, left=1.5in, right=1.5in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[icelandic]{babel}
\usepackage[T1]{fontenc}
\usepackage[sc]{mathpazo}
\usepackage[parfill]{parskip}
\renewcommand{\baselinestretch}{1.2}
% Tables and lists
\usepackage{booktabs,tabularx}
\usepackage{multirow}
\usepackage{enumerate}
\usepackage{adjustbox}
\usepackage{multicol}
\usepackage{xcolor}
\usepackage{algpseudocode}
\usepackage{tikz}
\usepackage{nicefrac}
\usepackage{changepage}
\usetikzlibrary{arrows, positioning, calc, graphs}

% Math
\usepackage{amsmath, amsfonts, amssymb, amsthm}
% Graphics

\usepackage{graphicx}
\usepackage{tikz}
% Code environment
\usepackage{minted}
%\usepackage{bm}
%\usepackage{siunitx}
%\usepackage{animate}
%\usepackage{hyperref}
%\usepackage{movie15}
%\usepackage{multicol}
%\usepackage{changepage}
\title{Formal Languages and Computability 4}
\author{Ragnar Björn Ingvarsson, rbi3}
\tikzset{->, >=stealth', shorten >=1pt, node distance=2cm,thick, main node/.style={circle,draw,minimum size=3em}}


\begin{document}
\renewcommand\thepage{}
	
	\maketitle

	\newpage
	\setcounter{page}{1}
	\renewcommand\thepage{\arabic{page}}
	\section{Give an example where $A$, $B$ are nonregular, $A\neq B$, but 
	$A\cap B$ is regular.}

	We can take for example the nonregular languages $A=\{0^n1^n | n\geq 0\}$
	 and $B=\{2^n3^n | n\geq0\}$ where $A\neq B$ and the intersection 
	 $A\cap B = \{\epsilon\}$ which is clearly regular.

	 \section{Let us define language $A$ accordingly, $A = \{i^nd^m | 0 
	 \leq n < m\}$. Show that $A$ is nonregular.}

	 We will use pumping lemma:

	 Let us assume that there exists a DFA that describes $A$, so that 
	 pumping lemma holds for some $p$. Let $s=i^pd^{p+1}$ so that 
	 $|s| = p+p+1 = 2p+1 > p$

	 Then we should be able to split $s$ into three pieces, $s = xyz$ so 
	 that $xy^iz\in A$ for all $i \geq 0$. Since $|xy| \leq p$, $x$ and $y$ 
	 both have to consist solely of $i$'s. Since that is the case, and 
	 $|y| > 0$, we can always find an $i$ so that $xy^iz$ gives us a case 
	 where $|xy^i| \geq |z|$, causing a contradiction.

	 Therefore, the language is nonregular.

	 \section{Let us consider ternary digits and a language $A$ which 
		 consists of strings $w$ that start with a prefix defining the most 
		 frequent symbol in the rest of the string. Show that no NFA exists 
		 that could recognize $A$}

		 We will use pumping lemma to prove that the language is nonregular. 
		 Let us assume that there exists a NFA that recognizes $A$ so that 
		 pumping lemma holds for some $p$. Let $s=02^p1^p0^p0$
		 so that $|s| = 3p+2$

		 We can then split any string $s$ into three parts, $s=xyz$ where 
		 $xy^iz\in A$ for all $i\geq0$. We see that since $|xy| \leq p$, 
		 there are a few cases for $y$:

		 \begin{itemize}
			 \item $y$ consists of only $0$. Here we can simply say that 
				 $i=0$ and get the string $2^p1^p0^p0$ where the string 
				 states that there are mostly $2$'s 
				 in the string which is incorrect and we get a contradiction.
			 \item $y$ consists of $0$ and then any number of $2$'s. 
				 Here we can repeat the same trick and get $xy^0z$, stating 
				 that there are mostly $2$'s in the string which is 
				 incorrect.
			 \item $y$ consists only of $2$'s. Here we can pump until $2$ 
				 is the most common symbol, overtaking the number of $0$'s, 
				 leading to a contradiction.
		 \end{itemize}

		 Here, all three cases can lead to a contradiction and therefore 
		 pumping lemma deduces that the language is nonregular and therefore 
		 there exists no NFA that could recognize it.


\end{document}
